\bf Excuse me, do you have a moment to talk about version control?

### 2017-05-30 README.tex
Re-render

Why would a statistician use version control, such as Git? And what is
the point of hosting your work online, e.g., on GitHub? Could the gains
possibly justify the inevitable pain?

@@

\begin{itemize}
\tightlist
\item
  Doing your work becomes tightly integrated with organizing, recording,
  and disseminating it. It's not a separate, burdensome task you are
  tempted to neglect.
\item
  Collaboration is much more structured, with powerful tools for
  asynchronous work and managing versions.
\item
  The marginal effort required to create a web presence for a project is
  negligible.
\item
  GitHub makes a fantastic course management system for courses that use
  R. You and your students can exchange actual working code and explore
  the associated results.
\item
  By using common mechanics across work modes (research, teaching,
  analysis), you achieve basic competence quickly and avoid the
  demoralizing forget-relearn cycle.
\end{itemize}

@@

In my opinion, \textbf{for new users}, the pros of Git only outweigh the
cons when you factor in the overhead of communicating and collaborating
with other people, including your future self. And who among us does not
need to do that? Your life is much easier if this is baked into your
workflow, as opposed to being a separate process that you dread and
avoid. Communication and collaboration are the killer apps of version
control. Git enforces a rigorous model of file management, but it is
critical for the distribution of files across different people,
computers, and time.

This has an implication for selecting your first Git projects: in order
to see substantial gain for your pain, you need to pick projects that
require you to share rapidly evolving files with others. It is tempting
to pick a quiet little project that you work on privately, but you risk
missing the main point of formal version control.

@@

The Git way is to track the evolution of that file, through a series of
commits, each equipped with an explanatory message and a nickname. All
collaborators sync regularly to a common version, acknowledging that the
difficulty of merging goes up faster than the size of the difference
\citep{good-enough}. Especially important versions get a human-readable
tag, to signal a meaningful milestone. Yes, there is some pain in
adopting the formalism of Git, but it is worth it.

@@

\href{https://github.com}{GitHub} is currently the most popular Git
hosting service. Others include \href{https://about.gitlab.com}{GitLab}
and \href{https://bitbucket.org}{Bitbucket}. These services provide a
home for Git-based projects on the internet. It is a bit like DropBox or
Google Drive, but more structured, powerful, and programmatic.

The remote host acts as the clearinghouse for a Git-managed project.
This allows other people to see the project files, sync up with the
current version, and perhaps even make or propose changes. These hosting
providers offer well-designed web interfaces that are a dramatic
improvement over traditional Unix Git servers. Many operations can be
done entirely in the browser, such as viewing current or past versions
of a file, commenting on a recent change, and editing or adding files.
These hosts also offer granular control over who can see, edit, and
administer a project.

@@

Here is the initial setup process, which you do once or, for some steps,
once per computer.

\begin{itemize}
\item
  Register for a free account with GitHub.
\item
  Install Git. Depending on your OS, Git might already be installed. But
  many of us may need to install it or will choose to update to a more
  recent version. Some basic configuration is critical, such as setting
  your username and email.
\item
  Install a local Git client \emph{(optional but highly recommended}). A
  Git client is software that provides a graphical user interface for
  Git, which is otherwise command-line only. If you are an R user, you
  will find that
  \href{https://www.rstudio.com/products/rstudio-desktop/}{RStudio}
  provides a great deal of this functionality. There are some notable
  gaps, however, so you might still choose to install a dedicated and
  comprehensive Git client such as
  \href{https://www.sourcetreeapp.com}{SourceTree} or
  \href{https://www.gitkraken.com}{GitKraken}. Git is a file-based
  system, so you can do some operations from RStudio, others from
  SourceTree, and others from the shell.
\item
  Confirm, with a practice repository, that local Git can send and
  receive the current version of the repository on GitHub, known as
  \textbf{pushing} and \textbf{pulling}, respectively. This will require
  authenticating yourself with GitHub from a local shell or Git client.
  At this point, most people elect to do a bit of extra setup to ensure
  that they are not repeatedly challenged for their GitHub credentials
  going forward.
\end{itemize}

Once this setup is done, you are ready to start using Git and GitHub
with your projects.

\subsection{Is this going to hurt?}\label{is-this-going-to-hurt}

Yes.

Git was built neither for the exact usage described here, nor for broad
usability. You will undoubtedly notice this, so it can be helpful to
know this in advance. Happily there are many helpful tools that mediate
your interactions with Git. GitHub itself is a fine example, as is
RStudio. In addition to pointing out tools that soften Git's sharpest
edges, I recommend specific habits and attitudes that reduce
frustration.

General recommendations for agony reduction:

@@

\textbf{Will it play especially well with Git/GitHub?} This boils down
to whether Git diffs will be informative and whether GitHub has nice
handling for the file type. Small-to-medium plain text files with hard
line breaks are a perfect fit, but there are a few more pleasant
surprises.

@@

\textbf{Will it cause problems with Git/GitHub?} This boils down to
whether Git diffs are informative, whether GitHub can render the file
usefully, and the file's likely effect on Git operations. A file that is
large and changing quite often can make your repository bloated and slow
down pushes and pulls. If a file is binary, such as a Word document or
Excel spreadsheet, you will not get human-readable diffs, nor can GitHub
display it in the browser. Html is not rendered on GitHub, but is,
instead, displayed in raw form. Binary files are also a prime source of
merge conflicts, because they are beyond the reach of Git's
sophisticated automatic merging logic. For many of us, adoption of
Git/GitHub should prompt a pivot away from \texttt{.docx} and
\texttt{.pdf} as primary output formats and towards \texttt{.md}, at
least during periods of rapid development.

@@

\emph{something I would like to write about, but have no space for here:
different models of collaboration, i.e.~everyone working on one repo \&
one branch vs.~making pull requests from branches in the main repo or
from forks}

@@

>>>>

Why would a statistician use version control, such as
\href{http://git-scm.com}{Git}? And what is the point of hosting your
work online, e.g., on \href{https://github.com}{GitHub}? Could the gains
possibly justify the inevitable pain?

@@

\begin{itemize}
\tightlist
\item
  Doing your work becomes tightly integrated with organizing, recording,
  and disseminating it. It's not a separate, burdensome task you are
  tempted to neglect.
\item
  Collaboration is much more structured, with powerful tools for
  asynchronous work and managing versions.
\item
  The marginal effort required to create a web presence for a project is
  negligible.
\item
  GitHub makes a fantastic course management system for courses that use
  R. You can exchange actual working code with your students and explore
  the associated results.
\item
  By using common mechanics across work modes (research, teaching,
  analysis), you achieve basic competence quickly and avoid the
  demoralizing forget-relearn cycle.
\end{itemize}

Now the bad news: Git was built neither for the exact usage described
here, nor for broad usability. You will undoubtedly notice this, so it's
best to know in advance. Happily, there are many helpful tools that
mediate your interactions with Git. GitHub itself is a fine example, as
is \href{https://www.rstudio.com/products/rstudio/}{RStudio}. In
addition to pointing out tools that soften Git's sharpest edges, I
recommend specific habits and attitudes that reduce frustration.

@@

In my opinion, \textbf{for new users}, the pros of Git only outweigh the
cons when you consider the overhead of working with other people,
including your future self. And who among us does not need to do that?
In a Git-based workflow, you document and, optionally, expose your work
as you go. Communication and collaboration are the killer apps of
version control. Git's model of file management can feel uncomfortably
rigid, but it enables the distribution of files across different people,
computers, and time.

This has an implication for selecting your first Git projects: you will
enjoy the most gain for your pain if you pick a project that involves
sharing rapidly evolving files with others. It is tempting to pick a
quiet, private project, but you risk missing out on the main benefits of
formal version control.

@@

The Git way is to track the evolution of that file, through a series of
commits, each equipped with an explanatory message and a nickname. All
collaborators sync regularly to a common version, acknowledging that the
difficulty of merging only goes up as more changes accumulate
\citep{good-enough}. Especially important versions get a human-readable
tag, to signal a meaningful milestone. Yes, there is some pain in
adopting the formalism of Git, but it is worth it.

@@

Git offers a powerful structure for file management and
\href{https://github.com}{GitHub} complements it with a slick user
interface and distribution mechanism for Git repositories. GitHub is
currently the most popular Git hosting service, providing a home for
Git-based projects on the internet. It is a bit like DropBox or Google
Drive, but more structured, powerful, and programmatic.

The remote host acts as the clearinghouse for a Git-managed project.
This allows others to see the project files, sync up with the current
version, and perhaps even propose or make changes. GitHub has a
well-designed web interface that is a dramatic improvement over
traditional Unix Git servers. Many operations can be done entirely in
the browser, such as viewing current or past versions of a file,
commenting on a recent change, and editing or adding files. GitHub also
offers granular control over who can see, edit, and administer a
project.

@@

Here is an overview of the initial setup process, which happens once or,
for some steps, once per computer. This is excerpted from
\href{http://happygitwithr.com}{Happy Git and GitHub for the useR},
which holds battle-tested instructions honed over several years in
\href{http://stat545.com}{STAT 545} at the University of British
Columbia.

\begin{itemize}
\item
  Register for a free account with GitHub.
\item
  Install Git. Depending on your OS, Git might already be installed. But
  many may need to install it or will choose to update to a more recent
  version. Some basic configuration is critical, such as setting your
  username and email.
\item
  Install a local Git client \emph{(optional but highly recommended}). A
  Git client is software that provides a graphical user interface for
  Git, which is otherwise command-line only. If you are an R user, you
  will find that
  \href{https://www.rstudio.com/products/rstudio-desktop/}{RStudio}
  provides a great deal of this functionality. There are some notable
  gaps, however, so you might still choose to install a dedicated and
  comprehensive Git client such as
  \href{https://www.sourcetreeapp.com}{SourceTree} or
  \href{https://www.gitkraken.com}{GitKraken}. Git is a file-based
  system, so you can do some operations from RStudio, others from
  SourceTree, and others from the shell.
\item
  Confirm, with a practice repository, that local Git can send and
  receive the current version of the repository on GitHub, known as
  \textbf{pushing} and \textbf{pulling}, respectively.
\end{itemize}

Once this setup is done, you are ready to start using Git and GitHub
with your projects. Some general recommendations for agony reduction:

@@

\textbf{Will it play nicely with Git/GitHub?} This boils down to whether
Git diffs will be informative and whether GitHub has nice handling for
the file type. Small-to-medium plain text files with hard line breaks
are ideal, but there are a few more pleasant surprises.

@@

\textbf{Will it cause problems with Git/GitHub?} This boils down to
whether Git diffs are informative, whether GitHub can render the file
usefully, and the file's likely effect on Git operations. A file that is
large and changing often can make your repository bloated and slow down
pushes and pulls. If a file is binary, such as a Word document or Excel
spreadsheet, you will not get human-readable diffs, nor can GitHub
display it in the browser. Html is not rendered on GitHub, but is,
instead, displayed in raw form. Binary files are also a prime source of
merge conflicts (see below), because they are beyond the reach of Git's
sophisticated automatic merging logic. This all implies that adoption of
Git/GitHub should prompt a pivot away from \texttt{.docx} and
\texttt{.pdf} as primary output formats and towards \texttt{.md}, at
least during periods of rapid development.

@@

@@

instructions re: supplementary materials
\url{http://amstat.tfjournals.com/supplementary-materials/}

TAS author instructions
\url{http://www.tandfonline.com/action/authorSubmission?journalCode=utas20\&page=instructions}
